Initial spatial conditions in simulation models: the missing leg of sensitivity analyses?

### 2017-08-04 paper.tex
.

\documentclass[Afour,sageh,times]{sagej}
\input{header.tex}

@@

% **************  MAIN BODY OF THE PAPER **************
\newpage

@@

@@

Comparing phase diagrams is useful because it allows us to compare the 
Indeed, we have as many phase diagrams than we have spatial grids, what makes a qualitative visual comparison not realistic\comment[FL]{donner un ordre de grandeur du nombre de simulations?}. A solution is to use systematic quantitative procedures. To our knowledge there exist no well established method to compare phase diagrams in agent-based modeling and geosimulation literature. We used a distance between two phase diagram corresponding to the share of between-diagrams variability relative to their internal variability\comment[FL]{je ne comprends pas ce choix : pourquoi faut il considérer que des classes ayant une forte cohérence interne seraient plus éloignées? ou alors la question ne se pose pas en ces termes... Par ailleurs est ce que la formule est bien homogène du point de vue des dimensions? En tous cas je pense qu'on peut se passer de la formule à ce stade, éventuellement à mettre dans l'affichage des résultats si on en a besoin.}

@@

%\begin{equation}\label{eq:phase-distance}
%d_r\left(\alpha_1,\alpha_2\right) = 2 \cdot \frac{d(f_{\vec{\alpha_1}},f_{\vec{\alpha_2}})^2}%{Var\left[f_{\vec{\alpha_1}}\right] + Var\left[f_{\vec{\alpha_2}}\right]}
%\end{equation}

@@

\subsection{Quantitative variation across grids}
\paragraph{Schelling} 
Within a standard Schelling model (i.e. initialised with a uniform density grid), \citet{Gauvinetal2009} have built the phase diagram of segregation patterns depending on the combination of parameter values. For high levels of tolerance ($S < 0.25$), there is no segregation. For high values of vacancies ($V > 0.65$) and low values of tolerance ($S > 0.5$), these is a diluted segregation state where homogeneous communities are separated from other by large empty buffers. Finally, for low values of vacancies ($V < 0.2$) and low values of tolerance ($S > 0.7$), the model is frozen in a state where everyone is unhappy but no-one can express their intolerant behaviour due to the lack of free spaces. Between these extreme cases, the model gives rise to segregated states where homogeneous communities adjoin one another. The objective of this quantitative experiment is to evaluate to which extent this phase diagram is modified when different density grids are applied. \comment[CC]{!!! Il faut retrouver les fichiers de simu !!!} \comment[CC]{Juste, tu penses qu'il faudrait faire la meme analyse que toi sur la distance entre phase diagrams? Ce serait cool mais si oui, j'aurais besoin d'aide (et des fichiers de simu!)}

\paragraph{Sugarscape} We measure the distance of all 3-dimensional phase diagrams to the reference phase diagram computed on the default model setup \comment[CC]{j'ai peut-etre manque une etape, mais c'est quoi le default model setup? en termes de grille d'entree: uniforme? random? ou Netlogo?} (see Fig.~\ref{fig:sugarscape-distance-meta} for its morphological positioning regarded generated grids), using equation~\ref{eq:phase-distance} with the L2 distance to ensure direct interpretability. \comment[CC]{il manque les equations} Indeed, it gives in that case the average squared distance between corresponding points of the phase diagrams, relative to the average of the variance of each. Therefore, values greater than 1 will mean that inter-diagram variability is more important than intra-diagram variability. \comment[CC]{en mode tatillon, on voudrait une double descente chromatique centree sur 1 pour la legende de la distance figure 4, qui distinguerait mieux les situations superieures a 1 et celle inf a 1 (ce qui nous interesse). Par exemple, blanc pour 1, et de ++ rose quand c'est >>1 et de ++ vert que c'est <<1. En R pour un level plot
}

@@

%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/relativedistance_morphspace}
\caption{\textbf{Relative distances of phase diagrams to the reference across grids.} Relative distance as a function of two first principal components of the morphological space (see text). Red point correspond to the reference spatial configuration. Green frame and blue frame give respectively the first and second particular phase diagrams shown in Fig.~\ref{fig:sugarscape-phasediagrams}.}
\label{fig:sugarscape-distance-pca}
\end{figure}
%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=0.24\textwidth]{figures/phasediagram_id27_maxSugar110}
\includegraphics[width=0.24\textwidth]{figures/phasediagram_id0_maxSugar110}
\caption{\textbf{Examples of phase diagrams.} We show two dimensional phase diagrams on $(P,s_-)$, both at fixed $s_+ = 110$. (Left) Green frame, obtained with $\alpha = 0.79$, $n=2$, $\beta = 0.14$, $N=157$; (Right) Blue frame, obtained with $\alpha = 2.56$, $n=3$, $\beta = 0.13$, $N=128$.}
\label{fig:sugarscape-phasediagrams}
\end{figure}
%%%%%%%%%%%%%

@@

>>>>

\documentclass[Royal,sageh,times]{sagej}
% Afour
\input{header.tex}

@@

% **************  MAIN BODY OF THE PAPER **************
%\newpage

@@

%\cite{ligmann2013spatially}

@@

Comparing phase diagrams is useful because it allows us to compare the 
Indeed, we have as many phase diagrams than we have spatial grids, what makes a qualitative visual comparison not realistic\comment[FL]{donner un ordre de grandeur du nombre de simulations?}. A solution is to use systematic quantitative procedures. To our knowledge there exist no well established method to compare phase diagrams in agent-based modeling and geosimulation literature. We used a distance between two phase diagram corresponding to the share of between-diagrams variability relative to their internal variability\comment[FL]{je ne comprends pas ce choix : pourquoi faut il considérer que des classes ayant une forte cohérence interne seraient plus éloignées? ou alors la question ne se pose pas en ces termes... Par ailleurs est ce que la formule est bien homogène du point de vue des dimensions? En tous cas je pense qu'on peut se passer de la formule à ce stade, éventuellement à mettre dans l'affichage des résultats si on en a besoin.}[(JR)]

@@

%\begin{equation}\label{eq:phase-distance}
%d_r\left(\alpha_1,\alpha_2\right) = 2 \cdot \frac{d(f_{\vec{\alpha_1}},f_{\vec{\alpha_2}})^2}{Var\left[f_{\vec{\alpha_1}}\right] + Var\left[f_{\vec{\alpha_2}}\right]}
%\end{equation}

@@

\subsection{Quantitative variation across grids}
\paragraph{Schelling} 
Within a standard Schelling model (i.e. initialised with a uniform density grid), \citet{Gauvinetal2009} have built the phase diagram of segregation patterns depending on the combination of parameter values. For high levels of tolerance ($S < 0.25$), there is no segregation. For high values of vacancies ($V > 0.65$) and low values of tolerance ($S > 0.5$), these is a diluted segregation state where homogeneous communities are separated from other by large empty buffers. Finally, for low values of vacancies ($V < 0.2$) and low values of tolerance ($S > 0.7$), the model is frozen in a state where everyone is unhappy but no-one can express their intolerant behaviour due to the lack of free spaces. Between these extreme cases, the model gives rise to segregated states where homogeneous communities adjoin one another. The objective of this quantitative experiment is to evaluate to which extent this phase diagram is modified when different density grids are applied.

\comment[CC]{Juste, tu penses qu'il faudrait faire la meme analyse que toi sur la distance entre phase diagrams? Ce serait cool mais si oui, j'aurais besoin d'aide (et des fichiers de simu!)}

\paragraph{Sugarscape} We measure the distance of all 3-dimensional phase diagrams to the reference phase diagram computed on the default model setup \comment[CC]{j'ai peut-etre manque une etape, mais c'est quoi le default model setup? en termes de grille d'entree: uniforme? random? ou Netlogo?}[(JR) c'est Netlogo, une configuration bi-centrique particulière] (see Fig.~\ref{fig:sugarscape-distance-meta} for its morphological positioning regarded generated grids), using equation~\ref{eq:phase-distance} with the L2 distance to ensure direct interpretability. \comment[CC]{il manque les equations} Indeed, it gives in that case the average squared distance between corresponding points of the phase diagrams, relative to the average of the variance of each. Therefore, values greater than 1 will mean that inter-diagram variability is more important than intra-diagram variability.

\comment[CC]{en mode tatillon, on voudrait une double descente chromatique centree sur 1 pour la legende de la distance figure 4, qui distinguerait mieux les situations superieures a 1 et celle inf a 1 (ce qui nous interesse). Par exemple, blanc pour 1, et de ++ rose quand c'est >>1 et de ++ vert que c'est <<1. En R pour un level plot}[(JR) oui tu as carrément raison, j'ai updaté les figures !]

@@

%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/relativedistance_morphspace}
\caption{\textbf{Relative distances of phase diagrams to the reference across grids.} Relative distance as a function of two first principal components of the morphological space (see text). Red point correspond to the reference spatial configuration. Green frame and blue frame give respectively the first and second particular phase diagrams shown in Fig.~\ref{fig:sugarscape-phasediagrams}.}
\label{fig:sugarscape-distance-pca}
\end{figure}
%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=0.48\textwidth]{figures/phasediagram_id27_maxSugar110}
\includegraphics[width=0.48\textwidth]{figures/phasediagram_id0_maxSugar110}
\caption{\textbf{Examples of phase diagrams.} We show two dimensional phase diagrams on $(P,s_-)$, both at fixed $s_+ = 110$. (Left) Green frame, obtained with $\alpha = 0.79$, $n=2$, $\beta = 0.14$, $N=157$; (Right) Blue frame, obtained with $\alpha = 2.56$, $n=3$, $\beta = 0.13$, $N=128$.}
\label{fig:sugarscape-phasediagrams}
\end{figure}
%%%%%%%%%%%%%

@@

Ref Clem : \cite{alizadeh2016generating} génération de réseau spatiaux / données synthétiques
